---
title: "Task 1 - Spatial data visualization of Oil Spill Incidents"
author: "Alex Ehrens"
date: "2/23/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# attach packages
library(tidyverse)
library(here)
library(janitor)
library(sf)
library(tmap)
library(gstat)
library(stars)
```

### Part 1. Read in spatial data and check/set projections
```{r}
# read in California oil spill incident dataset
oil_spills <- read_sf(here("task1", "Oil_Spill_Incident_Tracking_%5Bds394%5D-shp"), layer = "Oil_Spill_Incident_Tracking_%5Bds394%5D") %>% 
  clean_names()

# read in California county boundary dataset
ca_counties <- read_sf(here("task1", "ca_counties"), layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select(name)
```

##### Oil spill dataset projection:
```{r}
# check projections
st_crs(oil_spills)
```
##### California counties projection:
```{r}
st_crs(ca_counties)
```

Both shapefiles are already in the same (WGS_1984) projection, so there is no need to transform them into a new projection.

### Part 2 - Create an interactive map of oil spill locations in tmap
```{r}
# start with static exploratory map using ggplot
#ggplot() +
#  geom_sf(data = ca_counties) +
#  geom_sf(data = oil_spills)

# map an interactive map with tmap
tmap_mode("view")

tmap_options(basemaps = c("Esri.WorldTopoMap", "Esri.WorldGrayCanvas", "OpenStreetMap"))

tm_shape(oil_spills, name = "CA Oil Spills (2008)") +
  tm_dots("specificlo", 
          title = "Location of Oil Spill", 
          palette = c("mediumblue", "brown", "springgreen4")) 
```

### Part 3 - Create a static chloropleth of the count of oil spills by county
```{r}
# filter only inland oil spills
inland_oil_spills <- oil_spills %>% 
  filter(inlandmari == "Inland")

# join oil spill locations with counties dataset
spill_county_join <- ca_counties %>% 
  st_join(inland_oil_spills)

# get counts of oil spills by county
spill_counts <- spill_county_join %>% 
  count(name)
```

```{r}
# plot chloropleth of oil spill counts by county
ggplot(data = spill_counts) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgray","orange","red")) +
  theme_minimal() +
  labs(title = "Chloropleth of oil spill counts by county in California",
       subtitle = "Data from 2008",
       fill = "Number of oil \nspills by county",
       caption = "A chloropleth describing the spatial distribution and counts of oil spill events that occurred in each \nCalifornia county in 2008. Data from the CDFW Oil Spill Incident Tracking dataset.") +
  theme(plot.caption = element_text(hjust = 0))
```

### END TASK 1